---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
keep_md: true
---


## Loading and preprocessing the data

```{r Load_Data}
setwd("R:/RepData_PeerAssessment1")
activity <- read.csv(unz("activity.zip", "activity.csv"),stringsAsFactors=FALSE)
# Get sum of steps by day
sum_steps_by_day<-aggregate(activity[,1],list(Date=activity$date),sum,na.rm=TRUE)
colnames(sum_steps_by_day)[2] <- "Total_Steps"
# Get mean of steps by day
mean_steps_by_day<-aggregate(activity[,1],list(Date=activity$date),mean,na.rm=TRUE)
colnames(mean_steps_by_day)[2] <- "Mean_Steps"
# Get median of steps by day
median_steps_by_day<-aggregate(activity[,1],list(Date=activity$date),median,na.rm=TRUE)
colnames(median_steps_by_day)[2] <- "Median_Steps"
# Get mean of steps by 5-minute interval
mean_steps_by_interval<-aggregate(activity[,1],list(Interval=activity$interval),mean,na.rm=TRUE)
colnames(mean_steps_by_interval)[2] <- "Mean_Steps"
```

## What is mean total number of steps taken per day?

```{r Hist_Mean_Median}
# Generate histogram of steps by day
hist(sum_steps_by_day$Total_Steps,main="Steps Per Day",xlab="Steps",col="Red")
# Mean by day for steps
mean_steps_by_day
# Median by day for steps
median_steps_by_day
```


## What is the average daily activity pattern?

```{r Daily_Activity}

# Show average number of steps by time interval on a time series chart
plot(mean_steps_by_interval$Interval, mean_steps_by_interval$Mean_Steps, type="l", xlab= "Time Interval", ylab= "Average Number of Steps", col="black" , lwd=3)
# Which interval has the greatest average number of steps?
subset(mean_steps_by_interval, Mean_Steps == max(mean_steps_by_interval$Mean_Steps))
```

## Imputing missing values

For the missing values, use the mean value for the time interval (across all days)
as a proxy for the missing value.

```{r Missing_Data}
# Count the number of rows, dates/intervals, that have missing step count
sum(is.na(activity$steps))
# Replace NA values with the average for that time interval
activity2 <- activity
activity2$interval_step_mean <- mean_steps_by_interval$Mean_Steps[match(activity2$interval,mean_steps_by_interval$Interval)]
activity2$steps[is.na(activity2$steps)] <- activity2$interval_step_mean[is.na(activity2$steps)]

# Get sum of steps by day
sum_steps_by_day2<-aggregate(activity2[,1],list(Date=activity2$date),sum,na.rm=TRUE)
colnames(sum_steps_by_day2)[2] <- "Total_Steps"
# Get mean of steps by day
mean_steps_by_day2<-aggregate(activity2[,1],list(Date=activity2$date),mean,na.rm=TRUE)
colnames(mean_steps_by_day2)[2] <- "Mean_Steps"
# Get median of steps by day
median_steps_by_day2<-aggregate(activity2[,1],list(Date=activity2$date),median,na.rm=TRUE)
colnames(median_steps_by_day2)[2] <- "Median_Steps"
# Get mean of steps by 5-minute interval
mean_steps_by_interval2<-aggregate(activity2[,1],list(Interval=activity2$interval),mean,na.rm=TRUE)
colnames(mean_steps_by_interval2)[2] <- "Mean_Steps"
# Generate histogram of steps by day
hist(sum_steps_by_day2$Total_Steps,main="Steps Per Day (Imputed Missing Data)",xlab="Steps",col="Red")
# Mean by day for steps
mean_steps_by_day2
# Median by day for steps
median_steps_by_day2
```


## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable, wkday_end to designate Weekday or Weekend.
Saturday and Sunday are used for weekend days.

Compare average steps by interval weekend days vs. weekday days.

```{r Weekday_vs_Weekend}
activity2$wkday_end[weekdays(as.Date(activity2$date))=="Saturday" | weekdays(as.Date(activity2$date))=="Sunday"] <- "Weekend"

activity2$wkday_end[weekdays(as.Date(activity2$date))!="Saturday" & weekdays(as.Date(activity2$date))!="Sunday"] <- "Weekday"

mean_steps_by_interval_wkday_end<-aggregate(activity2[,1][activity2$wkday_end=="Weekday"],
                                     list(Interval=activity2$interval[activity2$wkday_end=="Weekday"]),
                                     mean,na.rm=TRUE)
colnames(mean_steps_by_interval_wkday_end)[2] <- "weekday"
mean_steps_by_interval2_we<-aggregate(activity2[,1][activity2$wkday_end=="Weekend"],
                                      list(Interval=activity2$interval[activity2$wkday_end=="Weekend"]),
                                      mean,na.rm=TRUE)

mean_steps_by_interval_wkday_end$weekend <- 
     mean_steps_by_interval2_we$x[match(mean_steps_by_interval_wkday_end$Interval,mean_steps_by_interval2_we$Interval)]

library(lattice)

xyplot(weekday + weekend ~ Interval, 
       data = mean_steps_by_interval_wkday_end, 
       layout = c(1,2), 
       type = "l", 
       ylab="Number of Steps",
       outer = TRUE)
```